---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r message=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5, tibble.max_extra_cols = 5)


library(geoclient)
```

# geoclient

[![Travis-CI Build Status](https://travis-ci.org/austensen/geoclient.svg?branch=master)](https://travis-ci.org/austensen/geoclient)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/austensen/geoclient?branch=master&svg=true)](https://ci.appveyor.com/project/austensen/geoclient)
[![Coverage Status](https://img.shields.io/codecov/c/github/austensen/geoclient/master.svg)](https://codecov.io/github/austensen/geoclient?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)


Tools to work with NYC's _Geoclient_ REST API.


This packages uses NYC's Geoclient API but is neither endorsed nor supported
by the the City of New York.

For information about the Geoclient API visit
[NYC's Developers Portal](https://developer.cityofnewyork.us/api/geoclient-api).


### Installation

Install from Github with [remotes](https://github.com/r-lib/remotes):

```{r eval=FALSE, include=TRUE}
# install.packages("remotes")
remotes::install_github("austensen/geoclient")
```


### Set up _Geoclient_ API keys

You can acquire your Geoclient app ID and Key by first registering with the
[NYC's Developer Portal](https://developer.cityofnewyork.us/user/register?destination=api)
at, then [create a new project](https://developer.cityofnewyork.us/create/project),
selecting "Geoclient v1" from available APIs. 

To avoid having to provide the ID and Key with each function call you can use `geoclient_api_keys()` to add your Geoclient app ID and Key to your `.Renviron` file
so they can be called securely without being stored in your code. 

```{r eval=FALSE, include=FALSE}
library(geoclient)

geoclient_api_keys("xxx", "xxx")
```


### Basic Usage

There are 6 main location types that can be set with _Geoclient_: Address, BBL (Borough-Block-Lot), BIN (Building Identification Number), Blockface, Intersection, and Place ("well-known NYC place name"). All of these functions return the results of the _Geoclient_ API call as a dataframe, with additional columns for the arguments provided to the function.

```{r}
geo_address(
  house_number = "139", 
  street = "MacDougal St", 
  borough = "MN",
  zip = "10012"
)
```

You can also pull out just a single column if that is all you need.

```{r}
df <- tibble::tribble(
  ~num,  ~st,                ~boro,         ~zip,
  "139", "MacDougal St",     "manhattan",   "11231",
  "295", "Lafayette street", NA,            "10012-2722",
  "40",  "WASHINGTON SQ S",  "MN",          NA
)

dplyr::mutate(df, bbl = geo_address(num, st, boro, zip)[["bbl"]])
```

For each of these location types there are two functions in this package that allow the arguments to be supplied either as individual vector, or with a dataframe and bare column names.

```{r}
geo_address_data(df, num, st, boro, zip)
```

The return dataframe will always be the same length and in the same order, so you can easily add all the return columns to your existing dataframe.

```{r}
dplyr::bind_cols(df, geo_address_data(df, num, st, boro, zip))
```

In addition to the 6 location types, _Geoclient_ also provides a single-field search option, which will guess the location type. This can be particularly helpful when you have address data that is not easily separated for use with `geo_address()`.

```{r}
df <- tibble::tribble(
  ~address,
  "139 MacDougal St manhattan, 11231",
  "295 Lafayette street, 10012-2722",
  "40 WASHINGTON SQ S MN"
)

geo_search_data(df, address)
```

